### 2023-05-10 docs/minutes/slides/document.tex
he-8 for earth model

% !TeX TS-program = xelatex

\documentclass[10pt]{beamer}

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{pgfplots}
\usepackage{xspace}

\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}
%\setsansfont[BoldFont={Fira Sans}]{Fira Sans Light}
%\setmonofont{Fira Mono}
%\usepackage[sfdefault]{Fira Sans}

\title{}
\subtitle{}

\author{}

\date{EE/GW meeting, May 11, 2023}

\begin{document}

	\maketitle

	\begin{frame}{}

	\begin{itemize}
		\item \alert{Problem = overly narrow,biased posteriors from nested sampling}. Even for ``easy" systems!
	\end{itemize}

Here is an easy system: $h = 10^{-2}, \sigma_m = 10^{-13}, \sigma_p = 0.0$ \newline 
\centering
\includegraphics[width=0.5\textwidth,height=0.5\textwidth]{images/example_system}

\raggedright
If the true declination $\delta = 1.0$, NS might give us a posterior of $\sim 0.95$ but variance v.v. small and posterior does not capture true value

\end{frame}

	\begin{frame}{}

	\begin{itemize}
		\item Why? Very nasty likelihood curves
	\end{itemize}

	$$  p(\theta | d) \propto p(d | \theta) \pi({\theta})$$
	$$ p(\theta | d) \propto \mathcal{L}(\theta) \pi({\theta})$$

	For uniform prior, posterior shape $\sim$ likelihood shape.

\end{frame}

\begin{frame}{}

	Example: $\mathcal{L}(\delta)$

\includegraphics[width=0.7\textwidth,height=0.6\textwidth]{images/likelihood_delta_1}

\end{frame}

\begin{frame}{}

	Example: $\mathcal{L}(\delta)$, zoomed 1

	\includegraphics[width=0.7\textwidth,height=0.6\textwidth]{images/likelihood_delta_2}

\end{frame}

\begin{frame}{}

	Example: $\mathcal{L}(\delta)$, zoomed 2

	\includegraphics[width=0.7\textwidth,height=0.6\textwidth]{images/likelihood_delta_3}

\end{frame}

\begin{frame}{}

	Example: $\mathcal{L}(\omega)$

	\includegraphics[width=0.7\textwidth,height=0.6\textwidth]{images/likelihood_omega_1}

\end{frame}

\begin{frame}{}

	Example: $\mathcal{L}(\omega)$, zoomed 1

	\includegraphics[width=0.7\textwidth,height=0.6\textwidth]{images/likelihood_omega_2}

\end{frame}

\begin{frame}{}

	Example: $\mathcal{L}(\omega)$, zoomed 2

	\includegraphics[width=0.7\textwidth,height=0.6\textwidth]{images/likelihood_omega_3}

\end{frame}

\begin{frame}{}

	All likelihood curves for $\bar{\theta}_{\rm GW}$:

	\includegraphics[width=0.9\textwidth,height=0.7\textwidth]{images/all_likelihood_curves}

\end{frame}

\begin{frame}{}

		\begin{itemize}
		\item \alert{Problem} = overly narrow,biased posteriors
		\item Cause: noisy, likelihoods which are "locally Gaussian". Sampling gets stuck in a local optima and infers posteriors widths based on that optima. 
%		\item Require $n_{live} > \frac{\rm prior \, volume}{\rm mode \,  footprint} = \frac{X}{\sigma^2}$
%		\begin{itemize}
%			\item For $\delta$ require $n_{live} > \frac{2 \pi}{10^{-5}} \sim 10^5 $
%			\item For $\omega$ require $n_{live} > \frac{10^4}{10^{-11}} \sim 10^7 $ 
%			\item Currently $n_{live} = 500$ parallelized on 32 cores $\sim 700$ minutes
%			\item Runtime scales $\mathcal{O}(n_{live}) \rightarrow 8000 $ hours for required $\delta$ sampling (just under 1 year!)
%
%		\end{itemize}
	\end{itemize}

\end{frame}

\begin{frame}{}
	Potential solutions:

	\begin{enumerate}
		\item Better settings for sampler? E.g. more live points (how many?), massively parallel, optimized likelihood evaluations,...
		\item Maximum likelihood / loss optimisation? 
		\item Drop PSR terms from model?
		\item MCMC at higher temperatures (trades off evidence..., less efficient	)
		\item other...?
	\end{enumerate}

\end{frame}

\begin{frame}{Drop PSR terms from model?}
	 \begin{equation}
		f_{\rm measured} = f_{\rm emitted} \, g(\tau; \bar{\theta})
		\label{eq:measureent}
	\end{equation}
	with
	\begin{align}
		g(\tau; \bar{\theta}) 
		& = 1 - \frac{1}{2} \frac{ H_{ij}q^i q^j }{(1 + \bar{n}\cdot \bar{q}) } \left[ \cos(-\Omega \tau +\Phi_0) - \cos(-\Omega \tau +\Phi_0 + \Omega (1 + \bar{n}\cdot \bar{q})  d) \right]
	\end{align}

Generate the fake data using the full expression for $g(\tau; \bar{\theta})$, but use $g'(\tau; \bar{\theta})$ in the Kalman measurement model

		\begin{align}
		g'(\tau; \bar{\theta}) 
		& = 1 - \frac{1}{2} \frac{ H_{ij}q^i q^j }{(1 + \bar{n}\cdot \bar{q}) } \left[ \cos(-\Omega \tau +\Phi_0) \right]
	\end{align}
\end{frame}

\begin{frame}{Drop PSR terms from model?}

		All likelihood curves for $\bar{\theta}_{\rm GW}$:
		\includegraphics[width=0.9\textwidth,height=0.7\textwidth]{images/all_likelihood_curves_psr}

\end{frame}

\begin{frame}{Option 2: Drop PSR terms from model?}

	\includegraphics[width=0.9\textwidth,height=0.7\textwidth]{images/canonical_mixed_result}

	OK, but some biases. c.f. location of the maxima on the previous slide
\end{frame}

%	\begin{frame}{}
	%
	%Numerical weather and climate modelling 
	%
	%        \begin{itemize}
		%	        \item Reduced precision floats + stochastic rounding
		%	        \item Automatic Differentiation
		%        \end{itemize}
	%    
	%    \alert{Open call for collaboration!}
	%    
	%    
	%	\end{frame}

\end{document}